---
title: "9.Model formation"
output: html_notebook
---

# 3 steps in regression model develop
1) model formation: all candidate explanatory variables and higher order term or interaction term.
2) variable selection: compare models to determine the best model 
3) model diagnostics: 
##residual analysis: 
##Influence statistcis: 

Model formation and varibale selection
```{r}
sat <- read.csv('desktop/sat.txt')
```
# pairspannel plot-see relationship between variable 
## raw scatterplot pairs
```{r}
source("desktop/pairs.panels.r")
pairs.panels(sat[,2:8])
```
## log scatterplot pairs
```{r}
sat2 <- data.frame(sat[,1:2], logtak=log(sat[,3]), sat[,4:8])
pairs.panels(sat2[,2:8])
```

# assess model fit: R-adj(big), AIG BIG Cp Press
after form model with all possible varibales, check if any variable need to be delete
## automatically check all subset model
### Cp-leaps(x,y) 
```{r}
library(leaps)
x <- sat2[,c(4,5)]
leaps(x, sat$sat)
```
Cp plot-choose the one with small cp value and near y=x line 
```{r}
mallow <- leaps(x, sat2$sat)
plot(mallow$size, mallow$Cp, main="Cp Plot", xlab="p", ylab="Cp", xlim=c(2,7), ylim=c(0,15),pch=19, cex=1.4)
abline(0,1,lwd=2)
```

### regsubsets
```{r}
subsets <- regsubsets(sat~income+years, data=sat2, nbest=2)
summary(subsets)
summary(subsets)$cp
summary(subsets)$adjr2
summary(subsets)$bic
```
1. nbest=1, just choose 1 best model for each number of parameter in model 
2. From the result we can see the result is the same with Cp leaps function, so these two are automatically caculate all possible combination of model, including interaction term.  

#### plot of subset-according to BIC and Cp 
```{r}
plot(subsets)
plot(subsets, scale="Cp")
```

### semiauto-step-AIC BIC(no interaction)
step(reg, k=2) AIC
step(reg2, k=log(n)) BIC show more panalty on model complexity, if complex BIC higher
#### backward elimination, each time delete one parameter 
if BIC n <- dim(sat2)[1] 
```{r}
null <- lm(sat~1, data=sat2)
full <- lm(sat~.-state, data=sat2)
bwd <- step(full, direction="backward", k=2)
```
#### forward elimination
```{r}
fwd <- step(null, scope=formula(full), direction="forward")
```
#### stepwise elimination
```{r}
stepwise <- step(null, scope=formula(full), direction="both")
```

## manually assess-same like step function
###1 
```{r}
reg1 <- lm(sat~logtak+income+years, data=sat2)
summary(reg1)
library(MPV)
PRESS(reg1)
step(reg1, k=2)
n <- dim(sat2)[1]
step(reg1, k=log(n))
```
###2 delete highest p-value income
```{r}
reg2 <- update(reg1,.~.-income)
summary(reg2)
PRESS(reg2)
step(...)
```
###3 delete highest p-value....until p-value are above 5%

notion: when use update to eliminate variable, need to pay attention to hierarch, when 2nd in,can't delete 1st order term. 






